---
title: "PSTAT 131 Homework 2"
author: "Sunrise Gao"
date: "10/14/2022"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## Linear Regression

For this lab, we will be working with a data set from the UCI (University of California, Irvine) Machine Learning repository ([see website here](http://archive.ics.uci.edu/ml/datasets/Abalone)). The full data set consists of $4,177$ observations of abalone in Tasmania. (Fun fact: [Tasmania](https://en.wikipedia.org/wiki/Tasmania "Tasmania") supplies about $25\%$ of the yearly world abalone harvest.)

The age of an abalone is typically determined by cutting the shell open and counting the number of rings with a microscope. The purpose of this data set is to determine whether abalone age (**number of rings + 1.5**) can be accurately predicted using other, easier-to-obtain information about the abalone.

The full abalone data set is located in the `\data` subdirectory. Read it into *R* using `read_csv()`. Take a moment to read through the codebook (`abalone_codebook.txt`) and familiarize yourself with the variable definitions.

Make sure you load the `tidyverse` and `tidymodels`!

```{r}

library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(ggthemes)
tidymodels_prefer()
tidymodels_packages()


abalone <- read.csv("abalone.csv")
head(abalone)

```

### Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

Assess and describe the distribution of `age`.

```{r}
abalone_new <- abalone %>%
  mutate(age = rings + 1.5)

head(abalone_new)

```

```{r}

hist(abalone_new$age, xlab = "Age",breaks =30, main = "Histogram of Abalone's Age", col = 'orange')

```
_As the result we got above, we can see that the age of abalone is not distributed evenly. The shape of the distribution is skewed to the left with a clear mode around age of 11, and most of abalone's age are around 8 to 13 and there is an outlier at around 30._

### Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

```{r}
set.seed(2000)

abalone_split <- initial_split(abalone_new, prop = 0.80, strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)

head(abalone_train)
head(abalone_test)
```

### Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,
    -   `longest_shell` and `diameter`,
    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.

```{r}
abalone_recipe <- abalone_train %>%
  recipe(age ~ type +longest_shell + diameter + height + 
           whole_weight + shucked_weight + viscera_weight + 
           shell_weight) %>%
  step_dummy(all_nominal_predictors())  %>%
  step_interact(terms = ~ starts_with("type"):shucked_weight+
                  longest_shell:diameter+
                  shucked_weight:shell_weight) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())
```


### Question 4

Create and store a linear regression object using the `"lm"` engine.

```{r}
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

### Question 5

Now:

1.  set up an empty workflow,
2.  add the model you created in Question 4, and
3.  add the recipe that you created in Question 3.

```{r}
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)
```

### Question 6

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
#fitting the model
lm_fit <- fit(lm_wflow, abalone_train)
```


```{r}
#predicting
abalone_hypo <- data.frame(type="F",longest_shell = 0.50, 
                           diameter = 0.10,
                           height = 0.30,
                           whole_weight = 4, 
                           shucked_weight = 1, 
                           viscera_weight = 2,
                           shell_weight = 1)
predict(lm_fit,new_data = abalone_hypo)

```

### Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).

```{r}

abalone_metrics <- metric_set(rsq, rmse, mae)

```

2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).

```{r}

abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))


abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))
abalone_train_res %>% 
  head()

```

3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.

```{r}

abalone_metrics(abalone_train_res, truth = age, estimate = .pred) 

```
_The R^2 represents how well the regression model fits the observed data. As the we getting rsq = 0.5567203, the model is not fitting the observed data well._\
_The root mean squared error is the standard deviation of the residuals (prediction errors), As we are predicting the age of abalone, a 2.1412871 of rmse is a bit to big as a prediction error._ \
_mae: The mean absolute error indicates that the magnitude of difference between the prediction of an observation and the true value of that observation is 1.5456314._ 
